---
title: "Initial AOIs for Image Acquisition"
author: "Lyndon Estes"
date: "7/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Several AOIs over Ghana will be collected as the primary locations for targetting image acquisitions (initially Planet).  These are provided as geojson AOIs.

We also want to collect imagery for a set of grid cells that are not in Ghana. These are for our existing Q sites, which are places where we have already gathered reference maps that we would like to refine, as well as a number of N sites, which are randomly selected throughout the continent. Workers will get served with N sites during those times when they are waiting for new sites to come in from `cvml`. Q sites will always be in rotation. 

Rather than polygon AOIs, we will provide these as a list of cell numbers that can be found in the master grid geotiff.  

## Ghana AOIs

```{r, message=FALSE, warning=FALSE}
library(raster)
library(sf)
library(dplyr)
library(leaflet)
library(rmapaccuracy)

meridia_sites <- read_sf("../data/processed/ghanasites.kml")
africa <- read_sf("../data/external/africa_noisl_gcs.sqlite")
africa_alb <- read_sf("../data/external/africa_countries_alb.sqlite")

create_aoi <- function(x, increase) {
  pol <- as(extend(extent(x), increase), "SpatialPolygons")
  return(pol)
}

# 1 degree boxes
towns <- unique(meridia_sites$town)
expanders <- list(rep(0.12, 4),# Asamankese
                  c(0.05, 0.2, 0.25, 0.05),  # Akwamu
                  c(0.005, 0.12, 0.005, 0.12),  # Dadieso
                  rep(0.22, 4),  # Ejura 
                  c(0.2, 0.2, 0.1, 0.3))  # Kulpawn
# aoi_akwamu <- create_aoi(meridia_sites %>% filter(town == "Akwamu"), 
#                          increase = c(0, 0, 0.25, 0.25))


aois <- do.call(rbind, lapply(1:length(towns), function(x) {
  # print(x)
  m <- meridia_sites %>% filter(town == towns[x])  
  pbox <- create_aoi(m, increase = expanders[[x]])
  st_as_sf(pbox)
}))
st_crs(aois) <- st_crs(meridia_sites)

f <- "../data/processed/image_target_aois.geojson"
if(!file.exists(f)) write_sf(aois, dsn = f)

km2 <- sum(st_area(st_transform(aois, crs = st_crs(africa_alb))) / 10000) / 100
paste("Area of AOIs =", round(unname(km2)), "km^2")
```

Map sites
```{r, fig.align='center', out.width = '100%'}
m <- leaflet() %>% #addTiles() %>% 
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(data = africa %>% filter(country == "Ghana"), weight = 2, 
              fillColor = "transparent", color = "white") %>% 
  addPolygons(data = aois, weight = 2, fillColor = "transparent", 
              color = "black") %>% 
  addPolygons(data = meridia_sites, weight = 1, col = "red")  
m
```

## Q and F sites

We'll connect into the database to get a temporary table that is holding a list of sites that we will use to assess worker accuracy. 
```{r}
coninfo <- mapper_connect(user = pgupw$user, password = pgupw$password,
                          db.tester.name = "lestes", 
                          alt.root = "data/processed", 
                          host = "crowdmapper.org")

# consensus Q and F fields
qf_names <- unique(DBI::dbGetQuery(coninfo$con, 
                            "SELECT name from qaqcfields_fromoldmaps")$name)    

# Select and prepare old Q sites
q_names <- (tbl(coninfo$con, "kml_data") %>% filter(kml_type == "Q") %>% 
  select("name") %>% collect())$name
q_flds <- unique(DBI::dbGetQuery(coninfo$con, # old sites with fields
                                  "SELECT name from qaqcfields")$name)
q_noflds <- q_names[!q_names %in% q_flds]  # old sites without fields
qf_all_flds <- unique(c(qf_names, q_flds))  # sites w/fields, old + consensus
set.seed(1)  # select 10% of old sites without fields
q_noflds_ss <- q_noflds[sample(1:length(q_noflds), 
                               round(length(q_noflds) * .1))]
qf_all_nms <- unique(c(qf_all_flds, q_noflds_ss))  # join them to sites w/fields

# get the ids and coordinates of those names in master grid
qf_ids <- coninfo$con %>% tbl("master_grid") %>% 
  filter(name %in% qf_all_nms) %>% select(id, x, y, name) %>% collect()
```

## N sites

Now get N sites, using new selection approach that will filter by fwt (>=5), to get sites more likely to have cropland. This will be coded in `select_n_sites`. "fwt" should be passed as an argument.
```{r}
# usual code for reading first available line
# first_avail_line <- as.numeric(
#   (tbl(coninfo$con, "system_data") %>% filter(key == 'firstAvailLine') %>%
#      collect())$value)
first_avail_line <- 1  # but replace with 1
kml_batch_size <- 1500
fwt <- 5

# select 
sql <- paste0("select gid, id, x, y, name, fwts from master_grid where ", 
              "gid >= ", first_avail_line, " and gid <= ",  
              (first_avail_line + kml_batch_size - 1), " and fwts >= ", 
              fwt)
              # fwt, " and name not like 'GH%'")
n_ids <- data.table(DBI::dbGetQuery(coninfo$con, sql))
# xy_tabs[, unique(substr(name, 1, 2))]

# select 10 sites at random from the selected list
set.seed(1)
n_ids_ss <- n_ids[sample(1:.N, 10, replace = FALSE), ]
gcstr <- proj4string(as_Spatial(meridia_sites))
grid.poly <- point_to_gridpoly(xy = n_ids_ss, w = 0.005, NewCRSobj = gcstr, 
                               OldCRSobj = gcstr)
```

Checking the 10 randomly selected sites from the sub-sample shows that most have some agriculture in them. 
```{r, fig.align='center', out.width = '100%'}
i <- 10
m <- leaflet() %>% 
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons(data = grid.poly[i, ], weight = 3, fillColor = "transparent", 
              col = "red")  
m

```

## Combine N, F, Q sites
```{r}
# any(!qf_ids$name %in% n_ids$name))
qfn_ids <- rbind(qf_ids, n_ids[, .(id, x, y, name)])
qfn_ids <- qfn_ids[order(id)]

f <- "../data/processed/individual_sites_needing_images.csv"
if(!file.exists(f)) fwrite(qfn_ids, file = f, row.names = FALSE)
```
